---
title: "The Effect of Writing Longer Papers"
subtitle: "ECON 480 - Econometrics - Fall 2018"
author: "Ryan Safner"
date: "Dec 10, 2018"
output: 
  beamer_presentation:
    incremental: true 
    fig_caption: yes
    toc: true 
    slide_level: 3
    includes:
      in_header: template.tex
    keep_tex: no 
    latex_engine: xelatex #required for Fira Sans font to render properly 

classoption: aspectratio=169
---

```{r setup, include=FALSE}
# For making transparent ggplot2 graphs, from https://gist.github.com/cboettig/5600558 

# Set plotting to bw plot default, but with transparent background elements.  
# Note transparency requires the panel.background, plot.background, and device background all be set!
library(tidyverse)
library(knitr)
library(stargazer)

theme_set(theme_bw(base_size=12))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
             plot.background = element_rect(fill = "transparent", colour = NA))
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, dev.args=list(bg="transparent"))
```

```{r, load-data}
papers<-read.csv("paperlengthsregcsv.csv")
```

### Overview 

- As a student writing an empirical research paper, does writing a longer paper attain a higher grade?
- Simple OLS regression
- Sample data collected from previous classes with paper assignments 
- Strong positive effect: for every marginal page written, grades improve by about 1-2 points 
- Robust to different models 

### Theory

- Dilemma:
1. Longer papers might imply students work hard and write a good paper 
2. But students might also put in low quality filler hoping to inflate their grades 

### Data

- I collected data from 7 different classes at 2 different colleges where I assigned a longer term paper 

\columnsbegin

\column{0.33\textwidth}

```{r}
papers %>% 
  count(Class) %>%
  kable(.)
```

\column{0.33\textwidth}

```{r}
papers %>% 
  count(School) %>%
  kable(.)
```

\column{0.33\textwidth}

```{r}
papers %>% 
  count(Year) %>%
  kable(.)
```

\columnsend

### Variables Used 

\begin{table}
  \begin{tabular}{ll}
  Variable & Description \\ \toprule
  Grade & Grade on paper assignment (0-100)\\ \midrule
  Pages & Number of pages written\\ 
  Female & $=1$ if student is female, $=0$ if male\\
  Morning & $=1$ if class met in morning, $=0$ if afternoon\\
  Hood & $=1$ if class taught at Hood, $=0$ if other school\\
  Metrics & $=1$ if class was Econometrics, $=0$ if other class\\ \bottomrule 
  \end{tabular} 
\end{table}

### Summary Statistics

```{r, results="asis"}
source("nice.summary.R")
sumstats<-nice.summary(papers,Grade,Pages)
sumstats
```

### Data: Histogram of Pages Written

```{r, fig.height=3.5}
library("ggplot2")
h<-ggplot(papers, aes(x=Pages))+
  geom_histogram(color="black", fill="indianred",breaks=seq(0,24,2))+
  scale_x_continuous(breaks=seq(0,24,2))+xlab("Pages Written")+ylab("Number of Students")+theme_light()
h
```

```{r, echo=F}
papers$Female<-ifelse(papers$Sex=="F",1,0)
papers$Morning<-ifelse(papers$Time=="Morning",1,0)
papers$Hood<-ifelse(papers$School=="Hood",1,0)
papers$Metrics<-ifelse(papers$Class=="Econometrics",1,0)
```

### Data: Histogram of Paper Grades

```{r}
g<-ggplot(papers, aes(x=Grade))+
  geom_histogram(color="black", fill="indianred",breaks=seq(0,100,10))+
  scale_x_continuous(breaks=seq(0,100,10))+xlab("Paper Grade")+ylab("Number of Students")+theme_light()
g
```

### Scatterplot of Pages Written and Grade with Outliers (0's) Dropped

```{r}

papers0<-subset(papers, Grade>0)

s0<-ggplot(papers0, aes(x=Pages, y=Grade))+
  geom_jitter(color="blue")+
  geom_smooth(method="lm", color="red")+
  scale_x_continuous(breaks=seq(0,24,2))+
  scale_y_continuous(breaks=seq(0,100,10))+xlab("Pages Written")+ylab("Paper Grades")+theme_light()
s0
```

### Scatterplot For Econometrics Classes vs Non-Econometrics Classes

```{r}
papers$Metrics<-ifelse(papers$Class=="Econometrics",1,0)

s0+facet_grid(cols=vars(Metrics))
```

### Empirical Model

$$Grade_i=Length_i+Final_i+Female_i+Morning_i+Hood_i+Metrics_i\epsilon_i$$

- $Length$ is the most important variable we care about
- $Length$ probably endogenous, correlated with other Grade-determining factors
    - Why we included $Final$, $Female$, $Morning$, $Hood$, and $Metrics$
- We probably expect $Length$ to be positive and small 
- **Endogeneity**: unmeasured attributes (e.g. paper topic, student's diligence, conscienciousness, intelligence, etc) correlated with both number of `Length` and `Grade`

### Results

```{r, regs, results="asis"}
basicreg<-lm(Grade~Pages, data=papers)
no0reg<-lm(Grade~Pages, data=papers0)
finalreg<-lm(Grade~Pages+Final, data=papers0)
sexreg<-lm(Grade~Pages+Final+Female, data=papers0)
timereg<-lm(Grade~Pages+Final+Female+Morning, data=papers0)
hoodreg<-lm(Grade~Pages+Final+Female+Morning+Hood, data=papers0)
metricsreg<-lm(Grade~Pages+Final+Female+Morning+Hood+Metrics, data=papers0)

stargazer(basicreg,finalreg,sexreg,timereg,hoodreg,metricsreg, type="latex", header=FALSE, float=FALSE, column.labels = c("All", "No 0s", "No 0s", "No 0s", "No 0s", "No 0s", "No 0s"), font.size = "tiny", column.sep.width = "-1pt", omit.stat=c("F", "rsq"), dep.var.caption = "")
```

### Results II

- For each additional page written, paper grades increase by about about 1-2 points 
- Still probably strongly endogenous 
- Poor advice to just write long papers! 
